---
title: "EDA plot"
author: "Yue Han"
date: "11/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "../")
```

```{r package, include=FALSE}
library(vroom)
library(tidyverse)
```

First of all, we load in some data from `merged_data.csv.gz`.

```{r load in data, results=FALSE}
if (!exists("track_data")) {
  track_data <- vroom("data/merged_data.csv.gz")
}
colnames(track_data)
```

Then we filtered the data according by WR's route: we only keep the play that at
least one WR's route is "GO", "CROSS", "SCREEN", "OUT" or "POST".

```{r subset data based on route}
# we only want to keep play whose route includes go, cross, screen, out or post
# we need to be careful when filtering data since only WR has route info
route_wanted <- c("GO", "CROSS", "SCREEN", "OUT", "POST")

# there are several WR in each play
# the following choose plays if there is one WR's route meets our requirement
included_play_id <- track_data %>%
  dplyr::select(playId, gameId, route) %>%
  filter(route %in% route_wanted) %>%
  dplyr::select(-route) %>%
  unique()

filtered_play_1 <- track_data %>%
  inner_join(included_play_id, by = c("playId", "gameId"))

# the following choose plays if all WR's route meets our requirement
route_unwanted <- track_data %>%
  dplyr::select(route) %>%
  unique() %>%
  drop_na() %>%
  filter(!route %in% route_wanted) %>%
  pull()

excluded_play_id <- track_data %>%
  dplyr::select(playId, gameId, route) %>%
  filter(route %in% route_unwanted) %>%
  dplyr::select(-route) %>%
  unique() %>%
  mutate(if_drop = TRUE)

filtered_play_2 <- filtered_play_1
#   left_join(excluded_play_id, by = c("playId", "gameId")) %>%
#   filter(is.na(if_drop)) %>%
#   dplyr::select(-if_drop)
```

Then we filter the play by the `offensePlayResult`, whihc is the yards that 
offense team gained, excluding penalty yardage.

We try 50 yards as the benchmark. It turns out to be only one play that 
satisfied our criteria.

```{r subset data based on play result}
# first we filter complete pass
filtered_play_3 <- filtered_play_2 %>%
  filter(passResult == "C")

# display by gained yards
filtered_play_3 %>% 
  dplyr::select(playId, gameId,offensePlayResult) %>%
  unique() %>%
  arrange(desc(offensePlayResult))

# then we filter offense play result that is bigger than 50
filtered_play_4 <- filtered_play_3 %>%
  filter(offensePlayResult > 50)

# there is only one game that satisfy above requirement
filtered_play_4 %>%
  dplyr::select(playId, gameId,offensePlayResult) %>%
  unique()
```

```{r slightly modified ggfootball, include=FALSE}
ggfootball_for_eda <- function(df, aes) {

  # Create data frames to add lines to field, hash marks, and 10-yard labels
  five_yard_df <- data.frame(x = seq(from = 15, to = 105, by = 5))
  hash_df <- data.frame(x = 11:109)
  yard_labels_df <- data.frame(
    x = seq(from = 20, to = 100, by = 10),
    y = rep(x = 4, n = 9),
    digits = c(
      seq(from = 10, to = 50, by = 10),
      seq(from = 40, to = 10, by = -10)
    )
  )
  
    # Create data frame for annotations
  annotate_df <- df %>% 
    group_by(frameId) %>% 
    slice(1) %>% 
    mutate(down_suffix = case_when(
      down == 1 ~ "st",
      down == 2 ~ "nd",
      down == 3 ~ "rd",
      down == 4 ~ "th"),
      quarter_suffix = case_when(
        quarter == 1 ~ "st",
        quarter == 2 ~ "nd",
        quarter == 3 ~ "rd",
        quarter == 4 ~ "th"
      ),
      play_start = paste0(quarter, quarter_suffix, " Quarter, ", down,
                          down_suffix, " and ", yardsToGo,
                          " from the ", yardlineSide, " ", yardlineNumber),
      timing = paste0("Quarter: ", quarter,
                      ", Game clock: ", game_clock_abr),
      title_string = paste0("Week ", week, ", ", gameDate, " at ",
                            game_time, " | ", homeTeamAbbr, " vs. ",
                            visitorTeamAbbr),
      score = paste0(homeTeamAbbr, ": ", preSnapHomeScore, " | ",
                     visitorTeamAbbr, ": ", preSnapVisitorScore))
  
  text_size <- 3

  ggplot(data = df) +
    
    # Make middle of field green
    geom_rect(
      data = NULL,
      aes(xmin = 10, xmax = 110, ymin = 0, ymax = 53.3),
      fill = "green4", color = "black", alpha = 0.1
    ) +

    # Add endzones
    geom_rect(
      data = NULL,
      aes(xmin = 0, xmax = 10, ymin = 0, ymax = 53.3),
      fill = "darkslategray", color = "black", alpha = 0.1
    ) +
    geom_rect(
      data = NULL,
      aes(xmin = 110, xmax = 120, ymin = 0, ymax = 53.3),
      fill = "darkslategray", color = "black", alpha = 0.1
    ) +

    # Format gridlines, tick marks, tick labels, and border of plot window
    theme_bw() +
    theme(
      panel.grid.minor = element_blank(),
      panel.grid.major = element_blank(),
      panel.border = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      axis.title = element_blank(),
      text = element_text(size = 12),
      # , legend.position = "none" # Optional hiding of legend
    ) +

    # Add vertical lines at each 5-yard increment
    geom_segment(
      data = five_yard_df,
      mapping = aes(x = x, xend = x, y = -Inf, yend = 53.3),
      color = "white"
    ) +

    # Add hash marks to field
    geom_segment(
      data = hash_df,
      mapping = aes(x = x, xend = x, y = 0.5, yend = 1.5),
      color = "white"
    ) +
    geom_segment(
      data = hash_df,
      mapping = aes(x = x, xend = x, y = 51.8, yend = 52.8),
      color = "white"
    ) +

    # Add yard line labels to field
    geom_text(
      data = yard_labels_df,
      mapping = aes(x = x, y = y, label = digits),
      color = "white"
    ) +
    geom_text(
      data = yard_labels_df,
      mapping = aes(x = x, y = 53.3 - y, label = digits),
      color = "white", angle = 180
    ) +

    # Create a final solid black outline for the field
    geom_rect(
      data = NULL,
      aes(xmin = 10, xmax = 110, ymin = 0, ymax = 53.3),
      fill = NA, color = "black"
    ) +

    # Add points and change colors of points in scatterplot
    geom_point(data = df, mapping = aes) +
    scale_color_manual(values = c("black", "slategray1", "red")) +

    # Label legend
    labs(color = "Legend") +

    # Add x and y axis limits
    lims(x = c(0, 120), y = c(-1, 59)) +
    
    # Add geom_text labels
    geom_text(data = annotate_df, mapping = aes(x = 0, y = 55,
                                                label = play_start),
              color = "black", hjust = 0, size = text_size) +
    geom_text(data = annotate_df, mapping = aes(x = 120, y = 55, 
                                                label = score),
              color = "black", hjust = 1, size = text_size) +
    geom_text(data = annotate_df, mapping = aes(x = 60, y = 59,
                                                label = str_wrap(playDescription, 
                                                                 80)),
              color = "black", hjust = 0.5, lineheight = 0.8, size = text_size) +
    labs(title = annotate_df$title_string[1]) +
    theme(plot.title = element_text(hjust = 0.5))
}
```

We plot the track of each player and football in that play.

```{r}
play_id <- filtered_play_3 %>% 
  dplyr::select(playId, gameId,offensePlayResult) %>%
  unique() %>%
  arrange(desc(offensePlayResult)) %>%
  slice(1)

p <- filtered_play_4 %>%
  filter(playId == play_id$playId, gameId == play_id$gameId) %>%
  ggfootball_for_eda(
  aes(
    x = x,
    y = y,
    color = team_2,
    alpha = frameId
  )
) +
  scale_alpha(range = c(0, 1)) +
  guides(alpha = FALSE)
p
```

```{r, include=FALSE}
ggsave(filename = paste0("EDA/eda_plots/",
                         play_id$offensePlayResult,
                         "_yards.png"), 
       plot = p)
```
